\label{sec:offline}

The surveyed articles span the period from the 60's to present. In
this section some tendencies for offline optimization systems are
surveyed.

Most authors choose to present a mathematical model of the problem, in
the spirit of the one presented in section \ref{sec:model}, which
optimizes for one or more of the parameters mentioned in section
\ref{sec:theproblem}. Attempts to produce a closed-form solution for the
problem are seen often eg. \cite{36}, which presents a general model
that has resemblances with the model in section \ref{sec:dynamicmodel}.

For representation of network layout, the models are based on either
graph theory or some kind of cell transmission scheme or petri net
system. The petri net representation is very popular and some examples
of its use are \cite{12}, \cite{16} and \cite{petri}.

In a petri representation of a traffic network, spaces on links are
represented by cells, which can fit a single vehicle. When a cell is
occupied no other vehicle may enter. Vehicles progress through the
network by making transitions to adjacent cells. The stop-bars of
intersections can be represented in the petri net as a line of cells
blocking a link facing the intersection. When the phase begins the
stop-bar cells are enabled and vehicles may pass.  The petri net
representation fits well into the context of traffic networks due to
its representation of concurrency (spaces in a lane) and triggered
actions (cross on green light).

For the model to be solvable without an external evaluation tool, such
as TRANSYT, it must contain functional constraints which relate
eg. the proper offset for signals along an arterial to flow-specific
parameters such as saturation flow rate and the platoon dispersion
factor.  The model must also adjust link flows according to the signal
settings, that is, the model must be bilevel, as introduced in section
\ref{sec:bilevel}. An example of such a model is seen in \cite{33} where a
genetic algorithm finds optimal signal timing plans, considering user
response to signal changes. TRANSYT is used to obtain a fitness (MOE)
value and the Path Flow Estimator for determination of the stochastic
user equilibrium.

Considering a typical optimization formulation involving a common
cycle time, green splits and offsets, it is clear that - even with some
discretization of time - the search space is vast, especially when
considering some form of simulation for obtaining the objective
value. An example of a heuristic designed to cut down the search
space is ADESS (see \cite{26}). Most often the heuristics are embedded
into the search routines and operate, for instance, by using "common
sense" rules such as pruning short cycle lengths under high
saturation or by exploiting sensitivity knowledge of the problem
\cite{40}.  Another way to deal with this issue is to employ a
metaheuristic search, which is simply run until a result is
needed. Some examples are: \cite{1} (tabu search), \cite{42} (particle
swarm optimization) and genetic algorithms, which are the most popular
metaheuristics, by far see \cite{13}, \cite{33}, \cite{43}, \cite{7},
\cite{41}, \cite{31}, \cite{27}, \cite{2} and \cite{26}.

Systems which operate in this manner are mostly suited to offline use
considering that some reported running times are as high as several
hours, even for small networks.  Some authors eg. \cite{16} turn
such a system into an adaptive system by re-running the optimization
procedure every $K$ cycle but clearly this strategy is not possible
for systems with long run time requirements. Furthermore the next set
of signal timing settings may be quite different from the current one
and changing plans in an instant will cause transient sideeffects such
as malfunctioning green waves.

Much work focuses on optimizations for single intersections or
coordination of signals along an arterial. Gartner et al. \cite{9}
gives a walkthrough of the most promising progression schemes at the
time (PASSER-II, NO-STOP-1) and extends the MAXBAND approach from a
single-arterial bandwidth optimization into MULTIBAND, which optimizes
bandwidth along multiple, possibly intersecting, arterials. \cite{6}
is an example of the selection and optimization of an arterial - or
priority route - with subsequent network-wide optimization, taking the
arterial optimization as a constraint.  In the article \cite{24}
Heydecker propose how to use existing optimization methods for single
intersections in combination to achieve a network optimization. Though
Heydecker admits that, because of the decentralization, it will be
difficult to obtain coordination.
