\label{sec:adaptive_cooperation}

Adaptive traffic control systems aims to coordinate signal controlled
intersections so as to optimize some performance index eg. average
delay or number of stops (or a combination) but also to reduce the
need for constant supervision and tuning of intersections.

They do this by dynamically adjusting cycle times, phase sequences and
green splits according to detected as well as predicted traffic and
thereby reacting to those dynamic aspects of traffic, which cannot be
captured by the static optimization routines used to generate
time-of-day plans. Some authors (\cite{1}, \cite{44}, \cite{46},
\cite{scoot2004}) even skip or work around the conventional periodic
scheme based on a common cycle time and make direct assignments of
phases and allow phases to be skipped, as presented in section
\ref{sec:dynamicmodel}.

It is evident that the cycle time is crucial in optimization because,
for a congested network, increasing the cycle time will always cause a
throughput increase (there are always cars waiting to cross the
intersection). In the litterature the cycle time is often common to
all intersections under traffic control so that green waves (arterial
progression) can be produced. For direct assignment systems the
throughput is dependent on allocation to phases of consecutive green
time. Long cycle times lead to long phase durations, which allow a
steady flow of vehicles to pass and minimizes lost and interphase time
per time unit.

For large networks the enforcement of a common cycle time is
inappropriate, however. Consider a network which is so large that two
disjoint arterials exist. In this case it is unlikely that a common
cycle time will allow green waves to exist for both arterials. For
networks of this type (size) a direct phase assignment model might
provide the necessary flexibility. Another feature of considering very
large networks is the possibility of traffic redirection. If it is
detected - or predicted - that an arterial is, or will be, congested
under current flow conditions it is sensible to redirect some traffic
onto alternative routes.

In this section some in-depth discussions are given for three adaptive
systems, which do not rely on offline optimization as it was presented
in section \ref{sec:offline}. The systems are:

\begin{enumerate}
\item \textbf{RHODES} by Pitu Mirchandani and Larry Head presented in
\cite{44} - a hierarchical system for network-wide optimization.

\item A \textbf{Phase-by-Phase} optimization strategy by Michael
Shenoda and Randy Machemehl presented in \cite{1} - a system using the
metaheuristic tabu search for determining greens for isolated
intersections in a phase-by-phase manner.

\item \textbf{DOGS} by Danish Technical Traffic Solution (TTS)
evaluated in the danish article \cite{dogs}, which provides
criteria-based capacity increases along an arterial.

\end{enumerate}

The systems will be compared in the areas of \textit{prediction} and
\textit{optimization strategy}.

\subsection{RHODES}
\input{rhodes}

\subsection{Phase-by-Phase}
\input{phase-by-phase}

\subsection{DOGS}
\input{dogs}

\subsection{Comparison}

The three systems presented here have different scope and as such
cannot be compared directly. The RHODES system is the most general,
being prepared for network wide optimization, though it will perform
well along an arterial, as shown in the FHWA test case. The DOGS
system is specifically designed to increase capacity along an
arterial. It is very flexible but requires many input parameters, for
which there is no clear selection strategy. Finally the phase-by-phase
system performs advanced intersection control but has no inherent
strategy for coordination, it its current state when the article was
written.

Some comparisons can be made in the two main topics covered for each
system ie. \textit{prediction} and \textit{optimization}.

\subsubsection*{Prediction}

In this area RHODES is by far the most advanced system. The systems
used at the intersection and sector level control are similar but in
the detail level and scope and build on existing work. Predictions are
made from detector data from upstream intersections using turning
probabilities and speeds, which can be extracted from historical data.

The Phase-by-Phase system, in its current state during the writing of
the article, had "perfect" prediction ie. the prediction process was
directly linked to the traffic generation process of the
simulator. The recommendation from Shenoda and Machemel is to further
investigate the ARIMA time-series analysis framework to obtain
accurate prediction for arrivals, which they argue is of dynamic
nature.

DOGS has no prediction capabilities. By studying the material on the
Herlev implementation and through talks with TTS and DRD it has become
clear that there is a general assumption that traffic, which enters
the arterial in either end is assumed to pass through the
arterial. Turn-in and turn-out movements are thus expected to cancel
out. This assumption could prove to be cumbersome if DOGS was to
operate longer arterials with differing traffic patterns in various
regions.

\subsubsection*{Optimization}

DOGS has a clear advantage in optimization since, for each cycle, the
only decision to be made is whether to switch between traffic actuated
control and pretimed signal plans or to change to a lower or higher
capacity program. This can be done in constant time using constraints
such as the one presented earlier. DOGS has no explicitly stated
objective function, only a \textit{political} objective - capacity
increase with reduced service for minor roads, which can be assessed
in rough terms using video detectors.

Phase-by-Phase minimizes delay using tabu search by adjusting green
time proportions assigned to phases. The initial solution is obtained
from a proportional heuristic, using Websters results, and the 1-bit
neighborhood function redistributes the green time for a single phase
to the next.  The running time of the search algorithm can be set to
the maximum time before a decision must be implemented eg. until the
end of the current phase.

RHODES optimization in the middle level involves generation and
traversal of a conflict resolution tree in order to find the best path
ie. conflict resolution according to some MOE. The optimization is
made for predictions of traffic 200-300 seconds in the future. The
article on RHODES does not mention how often the middle level
optimization is performed, though it seems natural to reoptimize every
time new predictions become available or continuously, if they are
updated before the optimization completes.  The conflict resolution
timing plans are fed to the intersection level optimization, which
fine tunes the plans according to predicted arrival of individual
vehicles using a dynamic programming approach. The optimization is
valid for 45-60 seconds into the future and is rerun at the end of
each phase.
