\label{sec:theproblem}

Essential parameters to the performance of a signalized intersection
-- or network of intersections -- are cycle time, green split and the
offset related to the common cycle time.

The amount of consecutive green time, which can be distributed to the
phases of an intersection depends on the cycle time. This is why the
cycle time is very important to the chosen MOE. In \citet{41} Sun et
al. explain that 

\begin{quote}[...] minimizing delay leads to short cycle length while
minimizing stops indicates long cycle length.
\end{quote}

The reason is that a long cycle length may stop a whole group of
vehicles (also known as a platoon) and incur a large delay for them,
but the platoon of vehicles only experience a single stop. With short
cycle time the vehicles never have to wait long for a green light but
are on average stopped more frequently.

%% Most authors who consider coordination of multiple signals use and
%% optimize a cycle time which is common for all traffic signals since it
%% would otherwise become hard to coordinate signals in static
%% schedules. In a small network, if each signal operates on its own
%% cycle time, green waves would only work on a periodic basis when the
%% cycle time and offset parameters of all signals eclipse. Otherwise the
%% green wave would drift and either cut off platoons of vehicles or give
%% too much green time.

The cycle time of an intersection is one of the most important
parameters in a signal plan since it sets  lower and upper bounds on
the green times for the phases. This relates to the MOE as well as the
safety of a network since 1) if a red light is shown for too long some
motorists may start to ignore the signals and 2) if the signals
cycle too frequently there is an increased risk of collisions.

The green split, which is the green-time to cycle-time ratio for the
phases of the intersection, should be considered as an indication of
the amount of traffic expected from each road facing the
intersection. Major roads will be given the larger split and minor
roads a smaller split.

The offset for an intersection is used to accommodate green waves for
traffic travelling from intersection to intersection at a specified
pace. This parameter is mostly relevant when optimizing signal plans
for major-minor types of arterials where platoons travel (mostly)
along a path, which traverses a string of intersections. Some of the
most promising results for this specific type of optimization are from
MAXBAND and MULTIBAND, described in \citet{37}.

Almost all methods optimize some or all of these three parameters. A
less commonly optimized parameter is the \textit{phase sequence}
ie. the order in which phases should be green during a cycle. This is
mostly relevant when prioritizing certain types of traffic, eg. public
transport (see \citet{scoot2004}) or emergency vehicles, but also when
considering safety aspects.

The problem of choosing an MOE depends on, among other things:

\begin{itemize}
\item The type of traffic which should gain benefits: private and commercial vehicles, public transportation, pedestrians and cyclists.
\item The network: highway, rural, urban.
\item Political objectives: safety, priority to businesses, reduction of emmisions combined with more public transportation and \textit{green} buses, which run on alternative fuels.
\end{itemize}

Clearly there are conflicts of interest among this selection of
objectives. One example has already been mentioned. Other examples of
objective conflicts are:

\begin{itemize}
\item Minimizing delays for vehicles along an artery will cause longer waiting times for crossing pedestrians.
\item Prioritizing eg. public transport by skipping a phase will lower the performance for private transport. 

In the SCOOT system (see \citet{scoot2004}) a phase skipping approach
to bus prioritization was implemented and tested in a London
intersection. The buses enjoyed 4 sec less delay but the delay for
non-bus vehicles increased by 1 sec in total on average. The vehicles
on the roads with no bus traffic suffered up to 14 sec delay on
average, however.

\item Any optimization which improves the experience of traversing the network by vehicle will probably cause more traffic and thus increase emissions.

\end{itemize}

Most of the articles in the survey use the average delay as an
objective. The TRANSYT optimization package by Robertson (1969) uses
the performance index (same as MOE) as explained in \citet{26}:

\begin{quote}
The performance index is defined as the sum for all signal-controlled
traffic streams of a weighted linear combination of estimated delay
and number of stops per unit of time and is used to measure the overall
cost of traffic congestion associated with the traffic control plan.
\end{quote}

Thus TRANSYT faces the multi-objectivity but the weights must be
defined and maintained by the traffic engineers supervising the
system. This is not a trivial task and requires intimate knowledge of
the network as well as the optimization package. The TRANSYT manual
suggests that weights should be set so that 1 stop $\simeq$ 20 sec
delay. This is in line with the guidelines used by the Danish Road
Directorate (DRD).

In \citet{41} a true multi-objective optimization approach is
tested. Sun et al. use a Non-dominated Sorting Genetic Algorithm (NSGA
II) to find Pareto-optimal solutions for the minimization objectives
\textit{average delay} and \textit{number of stops} per unit of time.
